The interactions of rational, pragmatic agents\\ lead to efficient language structure and use

### 2019-05-11 paper/paper.Rmd

### Reference games
Zipf’s example of optimal speaker- and listener-languages took the form of a reference game. We adopt that formulation here, assuming these communication games as our basic unit of analysis. In this framework, speakers and listeners are aware of a set of objects $M$ (*meanings*) and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1). Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$. Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}. We consider a set of contexts $C$ with an associated prior $P(C)$. Each context $c\in C$ describes a different distribution over meanings e.g. $p(M|C=c_i) \neq p(M|C=c_j)$. Finally, we consider a set of communicative events $e \in E$ where $<u, m, c> = e$ is an utterance-meaning-context triple.\par

@@

An RSA *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using $\ell$ in a given context $c$. That is, a speaker defines $P_{speaker}(u|m, c; \ell)$. We will use $S(u|m, c; \ell)$ as short-hand throughout.  A *listener agent* defines a conditional distribution over meanings, mapping from utterances $U$ to meanings $M$ using $\ell$ in a given context $c$ (i.e. $L(m|u, c;\ell)$). Note that both speakers and listeners can induce joint distributions over utterance-meaning pairs, although, importantly, these distributions may differ:
$$P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)$$
$$P_{listener}(u, m| c; \ell) = L(m|u, c; \ell)p(u|c)$$

@@

We assume that speaker effort is related to the surprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterances costs as independent from context (i.e.. $p(u|c)p(c)=p(u)p(c)$).} -- intuitively, the number of bits needed to encode the utterance $u$. This particular formalization of speaker-cost is general enough to accommodate a range of cost instantiations, such as production difficulty via articulation effort, cognitive effort related to lexical access, or others [@BennettGoodman2015a].\par

@@

As an example, say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), differing in cost, and two languages ($\ell_1$ and $\ell_2$), describing different utterance-meaning mappings. In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$). By contrast, in language $\ell_2$, $u_1$ can only refer to $m_1$ and $u_2$ can only refer to $m_2$ ($[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$). While it is cheaper for a speaker to use $\ell_1$ (because speaking it is always lower cost), it is more difficult for a listener (because $u_1$ is ambiguous). Crucially, however, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par

@@

\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {SI available at: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}. Code will be available upon acceptance.}}} \vspace{1em}
\noindent

>>>>

### Reference games
Zipf’s example of optimal speaker- and listener-languages took the form of a reference game. We adopt that formulation here, assuming these communication games as our basic unit of analysis. In this framework, speakers and listeners are aware of a set of objects $M$ (*meanings*) and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1). Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$. Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015)).}. We consider a set of contexts $C$ with an associated prior $P(C)$. Each context $c\in C$ describes a different distribution over meanings e.g. $p(M|C=c_i) \neq p(M|C=c_j)$. Finally, we consider a set of communicative events $e \in E$ where $<u, m, c> = e$ is an utterance-meaning-context triple.\par

@@

An RSA *speaker agent* defines a conditional distribution over utterances, mapping from intended meanings $M$ to utterances $U$ using $\ell$ in a given context $c$. That is, a speaker defines $P_{speaker}(u|m, c; \ell)$. We will use $S(u|m, c; \ell)$ as short-hand throughout.  A *listener agent* defines a conditional distribution over meanings, mapping from utterances $U$ to meanings $M$ using $\ell$ in a given context $c$ (i.e. $L(m|u, c;\ell)$). Speakers and listeners can induce joint distributions over utterance-meaning pairs, although, these distributions may differ:
$$P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)$$
$$P_{listener}(u, m| c; \ell) = L(m|u, c; \ell)p(u|c)$$

@@

We assume that speaker effort is related to the surprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterance costs as independent from context (i.e.. $p(u|c)p(c)=p(u)p(c)$).} -- intuitively, the number of bits needed to encode the utterance $u$. This particular formalization of speaker-cost is general enough to accommodate a range of cost instantiations, such as production difficulty via articulation effort, cognitive effort related to lexical access, or others [@BennettGoodman2015a].\par

@@

As an example, say we have two objects ($m_1$ and $m_2$), two utterances ($u_1$ and $u_2$), differing in cost, and two languages ($\ell_1$ and $\ell_2$), describing different utterance-meaning mappings. In language $\ell_1$, the low-cost $u_1$ can be used to refer to both $m_1$ and $m_2$ ($[[u_1]]_{\ell_1} = \{m_1, m_2 \}$), but the high-cost $u_2$ cannot be used at all ($[[u_2]]_{\ell_1} = \emptyset$). By contrast, in language $\ell_2$, $u_1$ can only refer to $m_1$ and $u_2$ can only refer to $m_2$ ($[[u_1]]_{\ell_2} = \{m_1\}$ and $[[u_2]]_{\ell_2} = \{m_2 \}$). While it is cheaper for a speaker to use $\ell_1$ (because speaking it is always lower cost), it is more difficult for a listener (because $u_1$ is ambiguous). Crucially, if context is disambiguating then the speaker can use $u_1$ to refer to either $m_1$ or $m_2$ and $\ell_1$ should be preferred to $\ell_2$.\par

@@

\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {SI and simulations: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}, \href{https://github.com/benpeloquin7/zipf_principles}{\url{https://github.com/benpeloquin7/zipf_principles}}}}} \vspace{1em}
\noindent

### 2019-05-11 paper/paper.tex
de-anonymized for psycharxiv

% KM added 1/4/18 to allow control of blind submission

@@

Zipf's example of optimal speaker- and listener-languages took the form
of a reference game. We adopt that formulation here, assuming these
communication games as our basic unit of analysis. In this framework,
speakers and listeners are aware of a set of objects \(M\)
(\emph{meanings}) and are knowledgeable about the set of possible
signals \(U\) (\emph{utterances}) that can be used to refer to a given
meaning (see Figure 1). Utterances may have different relative costs,
operationalized via a prior over utterances \(P(U)\). Similarly,
meanings differ in the relative degree to which they need to be talked
about, operationalized as a prior over meanings
\(P(M)\)\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}.
We consider a set of contexts \(C\) with an associated prior \(P(C)\).
Each context \(c\in C\) describes a different distribution over meanings
e.g. \(p(M|C=c_i) \neq p(M|C=c_j)\). Finally, we consider a set of
communicative events \(e \in E\) where \(<u, m, c> = e\) is an
utterance-meaning-context triple.\par

@@

An RSA \emph{speaker agent} defines a conditional distribution over
utterances, mapping from intended meanings \(M\) to utterances \(U\)
using \(\ell\) in a given context \(c\). That is, a speaker defines
\(P_{speaker}(u|m, c; \ell)\). We will use \(S(u|m, c; \ell)\) as
short-hand throughout. A \emph{listener agent} defines a conditional
distribution over meanings, mapping from utterances \(U\) to meanings
\(M\) using \(\ell\) in a given context \(c\) (i.e. \(L(m|u, c;\ell)\)).
Note that both speakers and listeners can induce joint distributions
over utterance-meaning pairs, although, importantly, these distributions
may differ: \[P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)\]
\[P_{listener}(u, m| c; \ell) = L(m|u, c; \ell)p(u|c)\]

@@

We assume that speaker effort is related to the surprisal of an
utterance in a particular
context\footnote{In the current set of simulations we consider utterances costs as independent from context (i.e.. $p(u|c)p(c)=p(u)p(c)$).}
-- intuitively, the number of bits needed to encode the utterance \(u\).
This particular formalization of speaker-cost is general enough to
accommodate a range of cost instantiations, such as production
difficulty via articulation effort, cognitive effort related to lexical
access, or others (Bennett \& Goodman, 2018).\par

@@

Importantly, we assume that events \(e = <u, m, c>\) are sampled
according the to following generative model -- some context occurs in
the world with probability \(P(C=c)\). Within this context, an object
\(m\) occurs with probability \(p(m|c)\). The speaker attempts to refer
to that object by sampling from her conditional distribution
\(S(u|m, c; \ell)\) (i.e. \(e \sim p(c)p(m|c)S(u|m, c; \ell)\)). From
these ingredients it is possible to derive the following objective
between the speaker and listener distributions (see
\href{https://bit.ly/2RBSGcU}{SI 2.1} for complete derivation).

@@

As an example, say we have two objects (\(m_1\) and \(m_2\)), two
utterances (\(u_1\) and \(u_2\)), differing in cost, and two languages
(\(\ell_1\) and \(\ell_2\)), describing different utterance-meaning
mappings. In language \(\ell_1\), the low-cost \(u_1\) can be used to
refer to both \(m_1\) and \(m_2\)
(\([[u_1]]_{\ell_1} = \{m_1, m_2 \}\)), but the high-cost \(u_2\) cannot
be used at all (\([[u_2]]_{\ell_1} = \emptyset\)). By contrast, in
language \(\ell_2\), \(u_1\) can only refer to \(m_1\) and \(u_2\) can
only refer to \(m_2\) (\([[u_1]]_{\ell_2} = \{m_1\}\) and
\([[u_2]]_{\ell_2} = \{m_2 \}\)). While it is cheaper for a speaker to
use \(\ell_1\) (because speaking is always low-cost), it is more
difficult for a listener (because \(u_1\) is ambiguous). Crucially,
however, if context is disambiguating then the speaker can use \(u_1\)
to refer to either \(m_1\) or \(m_2\) and \(\ell_1\) should be preferred
to \(\ell_2\).\par

@@

\fbox{\parbox[b][][c]{7.3cm}{\centering {SI available at: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}. Code will be available upon acceptance.}}}
\vspace{1em} \noindent

>>>>

% KM added 1/4/18 to allow control of blind submission
\cogscifinalcopy

@@

Zipf's example of optimal speaker- and listener-languages took the form
of a reference game. We adopt that formulation here, assuming these
communication games as our basic unit of analysis. In this framework,
speakers and listeners are aware of a set of objects \(M\)
(\emph{meanings}) and are knowledgeable about the set of possible
signals \(U\) (\emph{utterances}) that can be used to refer to a given
meaning (see Figure 1). Utterances may have different relative costs,
operationalized via a prior over utterances \(P(U)\). Similarly,
meanings differ in the relative degree to which they need to be talked
about, operationalized as a prior over meanings
\(P(M)\)\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015)).}.
We consider a set of contexts \(C\) with an associated prior \(P(C)\).
Each context \(c\in C\) describes a different distribution over meanings
e.g. \(p(M|C=c_i) \neq p(M|C=c_j)\). Finally, we consider a set of
communicative events \(e \in E\) where \(<u, m, c> = e\) is an
utterance-meaning-context triple.\par

@@

An RSA \emph{speaker agent} defines a conditional distribution over
utterances, mapping from intended meanings \(M\) to utterances \(U\)
using \(\ell\) in a given context \(c\). That is, a speaker defines
\(P_{speaker}(u|m, c; \ell)\). We will use \(S(u|m, c; \ell)\) as
short-hand throughout. A \emph{listener agent} defines a conditional
distribution over meanings, mapping from utterances \(U\) to meanings
\(M\) using \(\ell\) in a given context \(c\) (i.e. \(L(m|u, c;\ell)\)).
Speakers and listeners can induce joint distributions over
utterance-meaning pairs, although, these distributions may differ:
\[P_{speaker}(u, m | c; \ell) = S(u|m, c; \ell)p(m|c)\]
\[P_{listener}(u, m| c; \ell) = L(m|u, c; \ell)p(u|c)\]

@@

We assume that speaker effort is related to the surprisal of an
utterance in a particular
context\footnote{In the current set of simulations we consider utterance costs as independent from context (i.e.. $p(u|c)p(c)=p(u)p(c)$).}
-- intuitively, the number of bits needed to encode the utterance \(u\).
This particular formalization of speaker-cost is general enough to
accommodate a range of cost instantiations, such as production
difficulty via articulation effort, cognitive effort related to lexical
access, or others (Bennett \& Goodman, 2018).\par

@@

Importantly, we assume that events \(e = <u, m, c>\) are sampled
according to the following generative model -- some context occurs in
the world with probability \(P(C=c)\). Within this context, an object
\(m\) occurs with probability \(p(m|c)\). The speaker attempts to refer
to that object by sampling from her conditional distribution
\(S(u|m, c; \ell)\) (i.e. \(e \sim p(c)p(m|c)S(u|m, c; \ell)\)). From
these ingredients it is possible to derive the following objective
between the speaker and listener distributions (see
\href{https://bit.ly/2RBSGcU}{SI 2.1} for complete derivation).

@@

As an example, say we have two objects (\(m_1\) and \(m_2\)), two
utterances (\(u_1\) and \(u_2\)), differing in cost, and two languages
(\(\ell_1\) and \(\ell_2\)), describing different utterance-meaning
mappings. In language \(\ell_1\), the low-cost \(u_1\) can be used to
refer to both \(m_1\) and \(m_2\)
(\([[u_1]]_{\ell_1} = \{m_1, m_2 \}\)), but the high-cost \(u_2\) cannot
be used at all (\([[u_2]]_{\ell_1} = \emptyset\)). By contrast, in
language \(\ell_2\), \(u_1\) can only refer to \(m_1\) and \(u_2\) can
only refer to \(m_2\) (\([[u_1]]_{\ell_2} = \{m_1\}\) and
\([[u_2]]_{\ell_2} = \{m_2 \}\)). While it is cheaper for a speaker to
use \(\ell_1\) (because speaking it is always lower cost), it is more
difficult for a listener (because \(u_1\) is ambiguous). Crucially, if
context is disambiguating then the speaker can use \(u_1\) to refer to
either \(m_1\) or \(m_2\) and \(\ell_1\) should be preferred to
\(\ell_2\).\par

@@

\fbox{\parbox[b][][c]{7.3cm}{\centering {SI and simulations: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}, \href{https://github.com/benpeloquin7/zipf_principles}{\url{https://github.com/benpeloquin7/zipf_principles}}}}}
\vspace{1em} \noindent
